GPU-based Adaptive Beamforming for Active Sonar Imaging

### 2012-05-07 ECUA2012/gpu_capon_ecua2012.tex
la2pdf commit - 2012:05:08	01:45:47	(Tue 8.May)

We have tested our \gls{GPU} implementation of the \gls{MVDR} beamformer on two experimental dataset from the 32 element Kongsberg Maritime HISAS1030 sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV}, and used in sidescan mode to image the 1500 dwt oil tanker wreck Holmengraa (\Fig{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The \gls{MVDR} image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

@@

The other dataset was a 55\,kpixels sectorscan test image. We used it to compare the runtimes of our \gls{GPU} implementation of \gls{MVDR}, and an optimized single-thread C and Matlab implementation (\Fig{benchmarks}). Processing times for a are shown for varying subarray lengths $L$ and varying amounts of temporal averaging $N_\text{avg}$. The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of performing roughly 1Tflops.

@@

Figure \ref{holmengraa} demonstrates the \gls{MVDR} beamformer's ability to suppress interference and achieve better detail resolution. Compared the \gls{DAS} beamformer, the ship's edges come out as sharper, and the shadows are less noisy. But while \gls{DAS} processed this 1.3\,Mpixel image in milliseconds, our Matlab implementation of \gls{MVDR} needed minutes.

As Figure \ref{benchmarks} illustrates, using a \gls{GPU} improves matters. If we look at the total runtimes for a subarray size of $L=16$ and $2N_\text{avg}+1=5$ pixels of temporal averaging, the \gls{GPU} implementation is able to process the test image in roughly 40\,ms. This translates to a processing throughput of 1.1\,Mpixels/s. For the same scenario the C implementation clocked in at roughly 3\,s and Matlab at 17\,s, which is 75 and 435 times more than the \gls{GPU} implementation, respectively.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. Pixels in the image can be processed independently, and this also partially applies to the steps within each pixel.

>>>>

We have tested our \gls{GPU} implementation of the \gls{MVDR} beamformer on two experimental datasets from the 32 element Kongsberg Maritime HISAS1030 sonar~\cite{Hansen2009}. The sonar was attached to the HUGIN \gls{AUV}, and used in sidescan mode to image the 1500 dwt oil tanker wreck Holmengraa (\Fig{holmengraa}). Holmengraa is roughly 68\,m long and 9\,m wide, and lies on a slanted seabed at 77\,m depth. The \gls{MVDR} image was created with the parameters $L=16$, $N_\text{avg}=3$, and $d=1\%$, which proved to be a reasonable selection for this scenario.

@@

The other dataset was a 55\,kpixels sectorscan test image. We used it to benchmark our \gls{MVDR} implementation, and compared the result with the runtimes of an optimized single-thread C and Matlab implementation (\Fig{benchmarks}). Processing times are shown for varying subarray lengths $L$ and varying amounts of temporal averaging $N_\text{avg}$. The test system was a quad-core Intel Xeon E5620, with 64Gb of \gls{RAM}, and an nVidia Quadro 6000 card. This is a \gls{GPU} with 448 \gls{CUDA} cores and 6Gb of onboard \gls{RAM}, capable of performing roughly 1Tflops.

@@

Figure \ref{holmengraa} demonstrates the \gls{MVDR} beamformer's ability to suppress interference and achieve better detail resolution. Compared the \gls{DAS} beamformer, the ship's edges come out as sharper, and the shadows are less noisy. But while \gls{DAS} processed this 1.3\,Mpixel image in milliseconds, our Matlab implementation of \gls{MVDR} beamformer needed minutes.

As Figure \ref{benchmarks} illustrates, using a \gls{GPU} improves matters. If we look at the total runtimes for a subarray size of $L=16$ and $2N_\text{avg}+1=5$ pixels of temporal averaging, the \gls{GPU} implementation is able to process the test image in roughly 40\,ms. This translates to a processing throughput of 1.1\,Mpixels/s. For the same scenario the C implementation clocked in at roughly 3\,s and Matlab at 17\,s, which is 75 and 435 times slower than the \gls{GPU} implementation, respectively.

@@

The \gls{MVDR} beamformer is an algorithm well suited for implementation on a \gls{GPU}. Each pixels in the image can be processed independently, and the various steps within each pixel .
